Bucket-Sorted Independent Sets for Algebraic Multigrid

### 2007-09-23 bsis_paper/paper.tex
Major editing through the end of Section 3.
git-svn-id: file:///home/rblake/svn/pamg/trunk@28 c7f3fe13-8c30-0410-9127-9a9c42fe009c

\begin{abstract}
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
The algebraic multigrid (AMG) method~\cite{Ruge1987,Brandt1982} is an
efficient numerical algorithm for finding the solution to numerous
problems, including linear systems $Ax = b$ arising from the
discretization of partial differential equations on structured and
unstructured meshes. In many cases, the computational complexity of
AMG is $O(n)$, where $n$ is the number of unknowns in the linear
system. The linear cost scaling property of AMG makes the method
attractive, especially for large problems.

AMG algorithms are generally divided into two phases: the setup phase
and the solve phase. Setup phase algorithms provide functionality for
constructing elements needed by the solve phase, such as coarse grids,
operators to transfer information between levels of the problem, and
the linear systems to be solved on each level. The solve phase applies
relaxation methods on the problems generated in the setup phase and
uses the transfer operators to move vectors between levels.

In this paper, we focus attention on coarse-grid selection algorithms
in the AMG setup phase. One class of coarse-grid selection algorithms,
independent set-based algorithms, is studied in detail. In
Section~\ref{sec:cgs}, we introduce the coarse-grid selection problems
and the goals of coarse-grid selection algorithms. The components of
independent set-based coarsening algorithms are discussed.

\section{Coarse-Grid Selection}
\label{sec:cgs}
In AMG, coarse-grid selection is the process of selecting the degrees
of freedom for the coarse level problem. In classical AMG, degrees of
freedom on a coarse level are a subset of the degrees of freedom on a
finer level. Smoothed
aggregation~\cite{Vanek1996,Vanek2001,brezina2005}, on the other hand,
forms a coarse-level degree of freedom by aggregating several
fine-level degrees of freedom. In this paper, we study coarsening for
classical AMG.

Given a set of fine-level degrees of freedom, $\Omega_k$, coarse-grid
selection for classical AMG selects a set of coarse-level degrees of
freedom $\Omega_{k+1} \subset \Omega_k$. Several approaches have been
developed for selecting the coarse grid, with independent set-based
approaches the most numerous. Independent set-based algorithms trace
their roots to the classical AMG coarsening algorithm, RS
coarsening~\cite{Ruge1987}. These methods are our focus and are
discussed in more detail in the next section.

@@

\subsection{Independent Set-Based Selection}
\label{sec:indep-set-based}
Many independent set-based coarsening algorithms form $\Omega_{k+1}$
using the following heuristics.
\begin{description}
\item[H1:] For each unknown $j$ that strongly influences $F$-point
  $i$, $j$ is either a $C$-point or strongly depends on a $C$-point
  $k$ that also strongly influences $i$.
\item[H2:] The set of $C$-points needs to form a maximal independent
  set in the reduced graph of $S$ such that no $C$-point strongly
  depends on another $C$-point.
\end{description}
H1 and H2 cannot, in general, be satisfied simultaneously, so H1 is
enforced by independent set-based methods, while H2 serves as a
guideline whose purpose is to encourage selection of small
$|\Omega_{k+1}|$.

Some independent set-based coarsening algorithms use a variant of H1,
called \hp, which is a relaxed version of H1.
\begin{description}
\item[\hp] Each $F$-point must be strongly influenced by at least one
  $C$-point.
\end{description}

Independent set-based coarse grid selection algorithms implement three
routines to select coarse grids: initialization, independent set
selection, and weight update. A general independent-set based
coarsening algorithm is shown in
Algorithm~\ref{alg:general-indep-set}.
\begin{lstlisting}[caption={Independent Set-Based Coarse Grid
Selection},label=alg:general-indep-set]
%\textsc{Indep-Set-Coarse-Grid-Selection}$(\Omega)$:%
$F \leftarrow \emptyset$
$C \leftarrow \emptyset$
$w \leftarrow\,$%\textsc{Initialize-Weights}$(\Omega)$%
while $\Omega \ne \emptyset$
   $D \leftarrow\,$%\textsc{Select-Indep-Set}$(w,\, \Omega)$%
   $(F,\, C,\, \Omega,\, w) \leftarrow\,$%\textsc{Update-Weights}$(F,\,C,\, \Omega,\, w)$%
end
\end{lstlisting}

Each vertex in the graph is assigned a weight based on its merit to
becoming a $C$-point. The largest weights are conferred to vertices
best suited to become $C$-points, where the quality of potential
$C$-points is quantified by \emph{strength of connection}
measures. The classical strength of connection measure is based
entirely on the relative sizes of off-diagonal nonzeros in $A$. In the
classical definition, the set of unknowns that unknown $i$ strongly
depends upon is defined as
\begin{equation}
\label{2:eqn:S_i}
S_i = \left\{j: i \ne j \textrm{ and }|a_{ij}| \ge \theta \max_{k \ne
i}|a_{ik}|\right\},
\end{equation}
where $a_{ij}$ is the nonzero in row $i$, column $j$ of matrix $A$ and
$0 < \theta \le 1$. The set of unknowns that $i$ strongly influences,
denoted $S_i^T$, is defined as the set of unknowns that strongly
depend on $i$: $S_i^T = \{j: i \in S_j\}$.

@@

In classical coarsening, the independent set $D$ consists of a single
vertex $i$ in each iteration, where $w_i \ge w_j,\, \forall j \in S_i
\cup S_i^T$. Forming independent sets with only a single vertex is
inefficient in parallel, so parallel independent set methods use
different initialization techniques designed to enable the selection
of larger independent sets. In these algorithms, vertex $i$ is
eligible to be added to the independent set $D$ if the following
condition is satisfied:
\begin{equation}
  \label{eq:D-selection}
  w_i > w_j,\, \forall j \in S_i \cup S_i^T.
\end{equation}
The research presented here focuses primarily on the independent set
criterion in~(\ref{eq:D-selection}).

The weight update routines used in independent set-based coarsening
algorithms are varied but primarily take one of two forms: routines
that decrease the values of weights and routines that increase the
value of weights. In all cases, a vertex with a large weight has high
fitness for becoming a $C$-point.

Many independent set-based coarsening algorithms have been
developed. Ruge-St\"uben coarsening, also called classical coarsening,
is introduced in~\cite{Ruge1987}. The first parallel independent
set-based coarsening algorithms appeared approximately ten years ago
and include RS3~\cite{Henson2002}, CLJP~\cite{Cleary1998}, BC-RS and
Falgout~\cite{Henson2002}, PMIS and HMIS~\cite{Sterck2006},
CLJP-c~\cite{alber-cljpc}, PMIS-c1 and PMIS-c2~\cite{alber-PCGS}, and
PMIS Greedy~\cite{Butler2006}, to name a few.

@@

Theory is developed in this section to prove coarse-grid selection
invariance in general independent set-based algorithms. The algorithms
considered thus far select independent sets using
(\ref{eq:D-selection}), meaning vertex $i$ is eligible to be in $D$ if
its weight is larger than the weights of vertices in its neighborhood,
which is defined as follows.
\begin{defn}
\label{def:neighborhood}
The \emph{neighborhood} of vertex $i$, denoted $\mathcal{N}_i$, is the
union of the sets containing the strong influences of $i$ and the
strong dependencies of $i$: $\mathcal{N}_i = S_i \cup S_i^T$.
\end{defn}
Algorithms that rely on different and larger neighborhoods, such as
\emph{distance-$d$ neighborhoods}, are conceivable.
\begin{defn}
The \emph{distance-$d$ neighborhood} of $i$, denoted
$\mathcal{N}_i^d$, is the set of vertices within $d$ hops of $i$ in
the symmetrized strength matrix, excluding $i$. That is,
$\mathcal{N}_i^d = \left[\bigcup_{j \in
\mathcal{N}_i^{d-1}}\left(\mathcal{N}_j \cup \{j\}\right)\cup
\mathcal{N}_i\right] \setminus \{i\}$, where $d > 0$ and
$\mathcal{N}_i^0 = \emptyset$.
\end{defn}

@@

When a vertex $i$ satisfies the generalized selection
condition~(\ref{5:eqn:general-selection}), no other $C$-point
assignments affect $w_i$. This is formalized below.
\begin{lem}
\label{5:lemma:C-lock}
If (\ref{5:eqn:general-selection}) is satisfied for vertex $i$, then
$i$ must be the next vertex in $\{i\} \cup \mathcal{N}_i^s$ to become
a $C$-point, regardless of any other selections and updates made in
the graph.
\end{lem}
\begin{proof}
The satisfaction of (\ref{5:eqn:general-selection}) means $w_i > w_j$
for all $j \in \mathcal{N}_i^s$. For $i$ to not become a $C$-point,
its weight must become smaller than some $j \in \mathcal{N}_i^s$. The
weight of $i$, however, is not decremented unless some $j \in
\mathcal{N}_i^s$ satisfies~(\ref{5:eqn:general-selection}) and becomes
a $C$-point, which is impossible until after $i$ is assigned to the
$C$-point set.
\end{proof}

To demonstrate all algorithms using the same selection neighborhood
and update rules select identical coarse grids, a proof with an
inductive argument is presented below. The base case is provided by
the first set of vertices satisfying the general selection conditions.
\begin{defn}
Let $D_0$ be the set of all vertices satisfying
(\ref{5:eqn:general-selection}) in the first iteration of coarse-grid
selection.
\end{defn}

Vertices in $D_0$ are destined to become $C$-points regardless of the
algorithm used to build coarse grids. In most independent set-based
algorithms developed so far, all $D_0$ vertices become $C$-points in
the first iteration. Any algorithm constructed, however, eventually
selects all $D_0$ vertices as $C$-points, as proven in the following
lemma.
\begin{lem}
\label{5:lemma:D0_invariant}
Given the same selection neighborhoods and update heuristics, all
algorithms select vertices in $D_0$ as $C$-points.
\end{lem}
\begin{proof}
The proof follows from Lemma~\ref{5:lemma:C-lock}. All vertices in
$D_0$ satisfy the selection condition~(\ref{5:eqn:general-selection}),
so the assignment of any other $C$-point has no effect on the weight
of a $D_0$ vertex.
\end{proof}

@@

\begin{thm}
\label{5:theorem:invariance}
All independent set-based coarse-grid selection algorithms given the
same initial weights and using the same selection neighborhood,
selection criterion based on~(\ref{5:eqn:general-selection}), and
weight update heuristics as described above select identical coarse
grids.
\end{thm}
\begin{proof}
Let $c$ be the vertices in $\mathcal{N}_i^{2s}$ satisfying the
conditions for $D$ in some arbitrary iteration. Suppose assigning
vertices $c_1 \subset c$ to the $C$-point set leads to $w_i > w_j$ for
all $j \in \mathcal{N}_i^s$. Also suppose assigning $c_2 \subset c$,
$c_1 \ne c_2$, to the $C$-point set leads to the existence of some $j
\in \mathcal{N}_i^s$ such that $w_j > w_k$ for all $k \in
\mathcal{N}_j^s$.

@@

\begin{rem}
CLJP and CLJP-c use the distance-one neighborhood as the selection
neighborhood, (\ref{5:eqn:D-conditions}) as the selection criterion,
and the weight update heuristics in
Algorithm~\ref{3:alg:CLJP-update-heuristics}. Given the same initial
weights, all algorithms based on the parameters utilized by CLJP and
CLJP-c select identical coarse grids.
\end{rem}

@@

\section{Bucket Sorted Independent Set Selection}
\label{5:sec:contribution}

\subsection{CLJP-c}
CLJP-c~\cite{alber-cljpc} colors the graph of $S$ before selecting
$C$-points, and the colors are used as one component of the vertex
weights. As a result, the structure of the coarse grids selected is
improved. More formally, the use of graph coloring provides the
following important result.
\begin{thm}
\label{5:theorem:CLJPc-neighbor-independence}
For all pairs of vertices $i$ and $j \in \mathcal{N}_i$ CLJP-c
guarantees $w_i \ne w_j$.
\end{thm}
\begin{proof}
Assume two adjacent vertices $i$ and $j$ have the same weight. That is,
$|S_i^T| = |S_j^T|$ and the weight augmentation provided through
coloring is the same for $i$ and $j$. The graph of $S$, however, is
colored such that $i$ and $j$ are different colors for all $j \in
\mathcal{N}_i$, so a contradiction is reached.
\end{proof}
Theorem~\ref{5:theorem:CLJPc-neighbor-independence} establishes that
all adjacent vertices have different weights in CLJP-c, which is not
guaranteed in CLJP, although is unlikely to occur. The following
corollaries are a result of
Theorem~\ref{5:theorem:CLJPc-neighbor-independence}.
\begin{cor}
\label{cor:same-weight-indep-set}
Any set of vertices in the graph of $S$ sharing the same weight are an
independent set.
\end{cor}
\begin{cor}
\label{cor:max-weight-indep-set}
The set of vertices with the largest weight in the graph of $S$ form
an independent set satisfying~\ref{5:eqn:D-conditions}. That is, each
vertex in that independent set has a uniquely maximal weight in its
neighborhood.
\end{cor}
Corollary~\ref{cor:same-weight-indep-set} states that independent sets
can be selected in the graph simply by selecting sets of vertices with
the same weight. Corollary~\ref{cor:max-weight-indep-set} refines this
observation to a subset of vertices guaranteed to satisfy the
selection criterion. In particular, it shows it is possible to build
the coarse grid by selecting vertices with the maximum weight,
updating weights, selecting the next set of vertices with maximum
weight, and so on. This approach is taken by the algorithm developed
in Section~\ref{5:sec:contribution}. It is proven in
Section~\ref{5:sec:cg-invariance} that despite the difference in
approach, the resulting coarse grid is the same as that selected by
CLJP-c.

\subsection{BSIS}
\label{sec:bsis}
In this section, new techniques for searching the graph of $S$ for new
$C$-points and subsequently updating the weights of remaining vertices
are developed. The new algorithm is named Bucket Sorted Independent
Sets (BSIS) to reflect the data structure used.

>>>>

\begin{abstract}
Independent set-based coarse-grid selection algorithms are defined by
policies for how to do weight initialization, independent set
selection, and weight update. In this paper, we develop theory
demonstrating that all algorithms employing identical policies produce
identical coarse grids, regardless of the implementation. This result
on coarse grid invariance is applied to produce a new coarse-grid
selection algorithm that is substantially more efficient than an
existing algorithm using the same policies. Experimental results
comparing two versions of the new algorithm and the older algorithm
are presented, followed by a discussion on possible approaches to
effectively parallelize the new methods.
\end{abstract}

@@

\section{Introduction}
\label{sec:intro}
The algebraic multigrid (AMG) method~\cite{Ruge1987,Brandt1982} is an
efficient numerical algorithm for finding the solution to numerous
problems, including linear systems of the form $Ax = b$ arising from
the discretization of partial differential equations on structured and
unstructured meshes. In many cases, the computational complexity of
AMG is $O(n)$, where $n$ is the number of unknowns in the linear
system. The linear cost scaling property of AMG makes the method
attractive, especially for large problems.

AMG algorithms execute in two phases: the setup phase and the solve
phase. Setup phase algorithms provide functionality for constructing
elements needed by the solve phase, such as coarse grids, operators to
transfer information between levels of the problem (i.e., restriction
and prolongation operators), and the linear systems to be solved on
each level. The solve phase applies inexpensive iterative methods,
called relaxation methods in the context of AMG, on the problems
generated by the setup phase and uses the transfer operators to move
vectors between levels.

In this paper, we focus attention on coarse-grid selection algorithms
in the AMG setup phase. The independent set-based class of coarsening
algorithms is studied in detail. In Section~\ref{sec:cgs}, we discuss
current coarse-grid selection algorithms. The components of
independent set-based coarsening algorithms are
discussed. Section~\ref{sec:invariance} contains theory based on
general independent set-based coarsening and the invariance of coarse
grids selected when the same selection policies are
used. Section~\ref{sec:bsis} introduces a new coarsening algorithm,
Bucket-Sorted Independent Sets (BSIS), that, from the theory in
Section~\ref{sec:invariance}, produces the same coarse grids as
another coarsening algorithm in a much more efficient
manner. Experimental results are reported to demonstrate the
performance gains in BSIS discussion is included to propose a method
for parallelization of BSIS. Conclusions are presented in
Section~\ref{sec:conclusions}.

\section{Coarse-Grid Selection}
\label{sec:cgs}
Coarse-grid selection is the process of selecting the degrees of
freedom for the coarse-level problem. Classical AMG uses nodal
coarsening in which coarse-level degrees of freedom are a subset of
the degrees of freedom on finer levels. Smoothed
aggregation~\cite{Vanek1996,Vanek2001,brezina2005}, on the other hand,
forms a coarse-level degree of freedom by aggregating several
fine-level degrees of freedom. In this paper, we study nodal
coarsening.

Given a set of fine-level degrees of freedom, $\Omega_k$, nodal
coarse-grid selection constructs a set of coarse-level degrees of
freedom $\Omega_{k+1} \subset \Omega_k$. Several approaches have been
developed for selecting the coarse grid, with independent set-based
approaches being the most numerous. Independent set-based algorithms
trace their roots to the classical AMG coarsening algorithm,
Ruge-St\"uben (RS) coarsening~\cite{Ruge1987}, also called classical
coarsening. These methods are our focus and are discussed in more
detail following a brief overview of other techniques used for nodal
coarsening.

@@

\subsection{Independent Set-Based Selection}
\label{sec:indep-set-based}
Many independent set-based coarsening algorithms form $\Omega_{k+1}$
using the following heuristics.
\begin{description}
\item[H1:] For each unknown $j$ that strongly influences $F$-point
  $i$, $j$ is either a $C$-point or strongly depends on a $C$-point
  $k$ that also strongly influences $i$.
\item[H2:] The set of $C$-points needs to form a maximal independent
  set in the reduced graph of $S$ such that no $C$-point strongly
  depends on another $C$-point.
\end{description}
H1 and H2 cannot, in general, be satisfied simultaneously, so H1 is
enforced by independent set-based methods, while H2 serves as a
guideline whose purpose is to encourage selection of small
coarse-grid point sets (i.e., small $|\Omega_{k+1}|$).

Certain independent set-based coarsening algorithms use a variant of
H1, called \hp, which is designed to lead to sparser coarse-level
matrices.
\begin{description}
\item[\hp] Each $F$-point must be strongly influenced by at least one
  $C$-point.
\end{description}

Independent set-based coarse-grid selection algorithms implement three
routines to select coarse grids: initialization, independent set
selection, and weight update. A general independent-set based
coarsening algorithm is shown in
Algorithm~\ref{alg:general-indep-set}.
\begin{lstlisting}[caption={Independent Set-Based Coarse Grid
Selection},label=alg:general-indep-set]
%\textsc{Indep-Set-Coarse-Grid-Selection}$(\Omega)$:%
$F \leftarrow \emptyset$
$C \leftarrow \emptyset$
$w \leftarrow\,$%\textsc{Initialize-Weights}$(\Omega)$%
while $\Omega \ne \emptyset$
   $D \leftarrow\,$%\textsc{Select-Indep-Set}$(w,\, \Omega)$%
   $C \leftarrow C \cup D$
   $F \leftarrow F \setminus D$
   $(\Omega,\, w) \leftarrow\,$%\textsc{Update-Weights}$(D,\, \Omega,\, w)$%
end
\end{lstlisting}

Each vertex in the graph is assigned a weight based on its merit to
becoming a $C$-point. The largest weights are conferred to vertices
best suited to be $C$-points, where the quality of potential
$C$-points is quantified by \emph{strength of connection} measures. In
the classical definition, the set of unknowns that unknown $i$
strongly depends upon is based on the relative sizes of off-diagonal
nonzeros in row $i$ of $A$ and is defined as
\begin{equation}
\label{2:eqn:S_i}
S_i = \left\{j: i \ne j \textrm{ and }|a_{ij}| \ge \theta \max_{k \ne
i}|a_{ik}|\right\},
\end{equation}
where $a_{ij}$ is the nonzero in row $i$, column $j$ of matrix $A$ and
$0 < \theta \le 1$. The set of unknowns that $i$ strongly influences,
denoted $S_i^T$, is defined as the set of unknowns that strongly
depend on $i$: $S_i^T = \{j: i \in S_j\}$. The strong influence and
strong dependence sets are used in the definition of the
\emph{strength matrix} $S$, where the entry in row $i$, column $j$ of
$S$ is defined as
\begin{equation}
S_{ij} = \left\{
\begin{array}{ll}
  1 & \textrm{if } j \in S_i, \\
  0 & \textrm{otherwise}.
\end{array}\right.
\end{equation}

@@

The selection of the independent set $D$ works by comparing weights in
each node's neighborhood, which is defined as follows.
\begin{defn}
\label{def:neighborhood}
The \emph{neighborhood} of vertex $i$, denoted $\mathcal{N}_i$, is the
union of the sets containing the strong influences of $i$ and the
strong dependencies of $i$: $\mathcal{N}_i = S_i \cup S_i^T$.
\end{defn}

In classical coarsening, $D$ contains a single vertex $i$ in each
iteration, where $w_i \ge w_j,\, \forall j \in
\mathcal{N}_i$. Coarsening with independent sets containing a single
vertex per iteration is not scalable, so parallel independent set
methods use different initialization policies designed to enable the
selection of larger independent sets. In these algorithms, vertex $i$
is eligible to be added to the independent set $D$ if the following
condition is satisfied:
\begin{equation}
  \label{eq:D-selection}
  w_i > w_j\, \textrm{for all } j \in \mathcal{N}_i.
\end{equation}

The weight update routines in independent set-based coarsening
algorithms are varied but take one of two forms: routines that
decrease the values of weights and routines that increase the value of
weights. In all cases, a vertex with a large weight has high fitness
for becoming a $C$-point.

Many independent set-based coarsening algorithms have been
developed. RS coarsening is introduced in~\cite{Ruge1987} and is the
inspiration for many algorithms that followed. The first parallel
independent set-based coarsening algorithms appeared approximately ten
years ago and include RS3~\cite{Henson2002}, CLJP~\cite{Cleary1998},
BC-RS and Falgout~\cite{Henson2002}, PMIS and HMIS~\cite{Sterck2006},
CLJP-c~\cite{alber-cljpc}, PMIS-c1 and PMIS-c2~\cite{alber-PCGS}, and
PMIS Greedy~\cite{Butler2006}, to name a few.

@@

Theory is developed in this section to prove coarse-grid selection
invariance in general independent set-based algorithms. The algorithms
considered thus far select independent sets using
(\ref{eq:D-selection}), meaning vertex $i$ is eligible to be in $D$ if
its weight is larger than the weights of all vertices in its
neighborhood.

Algorithms relying on different and larger neighborhoods, such as
\emph{distance-$d$ neighborhoods}, are conceivable.
\begin{defn}
The \emph{distance-$d$ neighborhood} of $i$, denoted
$\mathcal{N}_i^d$, is the set of vertices within $d$ hops of $i$ in
the symmetrized strength matrix, excluding $i$. That is,
$\mathcal{N}_i^d = \left[\bigcup_{j \in
\mathcal{N}_i^{d-1}}\left(\mathcal{N}_j \cup \{j\}\right)\cup
\mathcal{N}_i\right] \setminus \{i\}$, where $d > 0$ and
$\mathcal{N}_i^0 = \emptyset$.
\end{defn}

@@

When a vertex $i$ satisfies the generalized selection
condition~(\ref{5:eqn:general-selection}), no other valid $C$-point
assignments are able to affect $w_i$. This is formalized below.
\begin{lem}
\label{5:lemma:C-lock}
If (\ref{5:eqn:general-selection}) is satisfied for vertex $i$, then
$i$ must be the next vertex in $\{i\} \cup \mathcal{N}_i^s$ to become
a $C$-point, regardless of any other selections and updates made in
the graph.
\end{lem}
\begin{proof}
The satisfaction of (\ref{5:eqn:general-selection}) means $w_i > w_j$
for all $j \in \mathcal{N}_i^s$. For $i$ to not become a $C$-point,
its weight must become smaller than some $j \in \mathcal{N}_i^s$. The
weight of $i$, however, is not decremented unless some $j \in
\mathcal{N}_i^s$ satisfies~(\ref{5:eqn:general-selection}) and becomes
a $C$-point, which is impossible until after $i$ is assigned to the
$C$-point set.
\end{proof}

To demonstrate all algorithms using the same selection neighborhood
and update rules select identical coarse grids, a proof with an
inductive argument is presented below. The base case is provided by
the first set of vertices satisfying the general selection conditions.

Let $D_0$ be the set of all vertices satisfying
(\ref{5:eqn:general-selection}) in the first iteration of coarse-grid
selection. Vertices in $D_0$ become $C$-points regardless of the
algorithm used to build coarse grids. In most independent set-based
algorithms developed so far, all $D_0$ vertices become $C$-points in
the first iteration. Any algorithm constructed, however, eventually
selects all $D_0$ vertices as $C$-points, as proven in the following
lemma.
\begin{lem}
\label{5:lemma:D0_invariant}
Given the same selection neighborhoods and update heuristics, all
algorithms select vertices in $D_0$ as $C$-points.
\end{lem}
\begin{proof}
The proof follows from Lemma~\ref{5:lemma:C-lock}. All vertices in
$D_0$ satisfy the selection condition~(\ref{5:eqn:general-selection}),
so the assignment of any other $C$-point has no effect on the weight
of a $D_0$ vertex.
\end{proof}

@@

\begin{thm}
\label{5:theorem:invariance}
All independent set-based coarse-grid selection algorithms given the
same initial weights and using the same selection neighborhood,
selection criterion based on~(\ref{5:eqn:general-selection}), and
weight update heuristics as described above select identical coarse
grids.
\end{thm}
\begin{proof}
Let $c$ be the vertices in the extended selection neighborhood of $i$
($\mathcal{N}_i^{2s}$) satisfying the conditions for $D$ in some
arbitrary iteration. Suppose assigning vertices $c_1 \subset c$ to the
$C$-point set leads to $w_i > w_j$ for all $j \in
\mathcal{N}_i^s$. Also suppose assigning $c_2 \subset c$, $c_1 \ne
c_2$, to the $C$-point set leads to the existence of some $j \in
\mathcal{N}_i^s$ such that $w_j > w_k$ for all $k \in
\mathcal{N}_j^s$. In the first case, $i$ becomes a $C$-point before
$j$, and the opposite is true in the second case.

@@

@@

\section{Bucket Sorted Independent Set Selection}
\label{sec:bsis}

\subsection{CLJP-c}
CLJP-c~\cite{alber-cljpc} colors the graph of $S$ before selecting
$C$-points, and the colors are used as one component of the vertex
weights. As a result, the structure of the coarse grids selected is
improved. More formally, the use of graph coloring provides the
following important result.
\begin{thm}
\label{5:theorem:CLJPc-neighbor-independence}
For all pairs of vertices $i$ and $j \in \mathcal{N}_i$ CLJP-c
guarantees $w_i \ne w_j$.
\end{thm}
\begin{proof}
Assume two adjacent vertices $i$ and $j$ have the same weight. That is,
$|S_i^T| = |S_j^T|$ and the weight augmentation provided through
coloring is the same for $i$ and $j$. The graph of $S$, however, is
colored such that $i$ and $j$ are different colors for all $j \in
\mathcal{N}_i$, so a contradiction is reached.
\end{proof}
Theorem~\ref{5:theorem:CLJPc-neighbor-independence} establishes that
all adjacent vertices have different weights in CLJP-c, which is not
guaranteed in CLJP, although is unlikely to occur. The following
corollaries are a result of
Theorem~\ref{5:theorem:CLJPc-neighbor-independence}.
\begin{cor}
\label{cor:same-weight-indep-set}
Any set of vertices in the graph of $S$ sharing the same weight are an
independent set.
\end{cor}
\begin{cor}
\label{cor:max-weight-indep-set}
The set of vertices with the largest weight in the graph of $S$ form
an independent set satisfying~\ref{5:eqn:D-conditions}. That is, each
vertex in that independent set has a uniquely maximal weight in its
neighborhood.
\end{cor}
Corollary~\ref{cor:same-weight-indep-set} states that independent sets
can be selected in the graph simply by selecting sets of vertices with
the same weight. Corollary~\ref{cor:max-weight-indep-set} refines this
observation to a subset of vertices guaranteed to satisfy the
selection criterion. In particular, it shows it is possible to build
the coarse grid by selecting vertices with the maximum weight,
updating weights, selecting the next set of vertices with maximum
weight, and so on. This approach is taken by the algorithm developed
in Section~\ref{sec:bsissub}. It is proven in
Section~\ref{5:sec:cg-invariance} that despite the difference in
approach, the resulting coarse grid is the same as that selected by
CLJP-c.

\subsection{BSIS}
\label{sec:bsissub}
In this section, new techniques for searching the graph of $S$ for new
$C$-points and subsequently updating the weights of remaining vertices
are developed. The new algorithm is named Bucket Sorted Independent
Sets (BSIS) to reflect the data structure used.
